\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{geometry}
\usepackage{amsmath,amssymb}
\usepackage{siunitx}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage[round]{natbib}
\usepackage{hyperref}
\usepackage{xcolor}

\geometry{letterpaper, left=1in, right=1in, top=1in, bottom=1in}
\hypersetup{hidelinks=true}

\title{Uniform vertical regridding of radiosonde and dropsonde profiles}
\author{Thomas D. DeWitt}
\date{\today}

\begin{document}
\maketitle

\section{Introduction}

Radiosonde and dropsonde profiles report atmospheric variables at irregular vertical intervals that differ between instruments, processing methods, and field campaigns.  To compare profiles across datasets or to compute structure functions at well-defined separation distances, we first place all profiles on a common vertical grid.  The algorithm described here does so by constructing a uniform grid along the altitude axis and averaging all measurements that fall within each cell. If no measurements fall within a given cell, it is marked as missing.

The simplicity of this approach is deliberate. If interpolation was used, artificial smoothness would be introduced where data are interpolated given that commonly-used interpolation schemes assume a smooth function with $H=1$ \citep[cf.][]{lovejoy2007}.

Table~\ref{tab:datasets} summarizes the datasets included in this archive.

\begin{table}[h]
\centering
\caption{Source datasets.  Profile counts reflect the number used after quality filtering.}
\label{tab:datasets}
\small
\begin{tabular}{lrlll}
\toprule
Dataset & Profiles & Period & Region & Citation \\
\midrule
IGRA              & (subsampled) & 2000--2025                & Global                     & \citet{durre2006} \\
JOANNE            & 1068         & Jan--Feb 2020              & Tropical N Atlantic        & \citet{george2021} \\
BEACH             & 976          & Aug--Sep 2024              & Tropical Atlantic          & \citet{gloeckner2025} \\
OTREC             & 648          & Aug--Oct 2019              & E Pacific, Caribbean       & \citet{vomel2021} \\
ACTIVATE          & 801          & Feb 2020 -- Jun 2022       & W North Atlantic           & \citet{vomel2023activate} \\
Hurricane         & 13{,}678     & 1996--2012                 & Atlantic, E Pacific TCs    & \citet{wang2015} \\
DYNAMO            & 468          & Nov--Dec 2011              & Central Indian Ocean       & \citet{ciesielski2014} \\
SHOUT             & 638          & 2015--2016                 & Atlantic TCs, Pacific      & \citet{wick2020} \\
AR Recon          & ${\sim}$8000 & 2018--2026 winters         & E North Pacific            & \citet{cobb2022} \\
ENRR              & 976          & Jan--Mar 2016              & Tropical Pacific           & \citet{dole2018} \\
HS3               & 1505         & 2011--2014                 & Atlantic TCs               & \citet{braun2016} \\
PREDICT           & 558          & Aug--Sep 2010              & Caribbean, W Atlantic      & \citet{montgomery2012} \\
HALO--(AC)\textsuperscript{3}  & 463 & Mar--Apr 2022          & Arctic, Fram Strait        & \citet{ehrlich2025} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{IGRA}

The Integrated Global Radiosonde Archive \citep{durre2006} is a quality-assured compilation of radiosonde observations from more than 1500 globally distributed stations, with records extending from the 1960s to the present.  Observations typically include pressure, temperature, dewpoint depression, geopotential height, wind direction, and wind speed.  IGRA integrates data from disparate sources and applies automated quality assurance that removes measurements failing checks for physical implausiblity, internal inconsistency, climatological outliers, and temporal inconsistency.

We consider soundings from 2000--2025, subsampled to every 10th day (DOY 1, 11, 21, etc.) to reduce data volume.  The vertical grid spacing is \SI{100}{m} (compared to \SI{10}{m} for dropsonde datasets).  Each station is written to a separate file spanning whatever date range it has data for.  Station metadata (name, location, elevation, equipment history) is extracted from the IGRA metadata catalog and stored as global attributes.

IGRA header records provide a nominal observation hour \texttt{HOUR} (valid range 0--23, or 99 for missing) and an actual release time \texttt{RELTIME} in HHMM format (9999 when missing).  The IGRA v2.2 format specification notes that ``the relationship between this hour and the release time varies by data provider, over time, and among stations,'' so we do not use \texttt{HOUR} as a time label.  We take \texttt{RELTIME} as \texttt{launch\_time} (falling back to \texttt{HOUR} only when \texttt{RELTIME} is missing, and shifting by one day when \texttt{RELTIME} places the release more than 12 hours after a synoptic \texttt{HOUR}, as happens for pre-midnight launches labelled 00~UTC).  RELTIME is reliable: across the 212,825 profiles for which we can cross-check it against the elapsed-time field \texttt{ETIME}, it agrees with \texttt{ETIME}[0] to within one minute for 99\% of profiles (median 0~s, 99th percentile 24~s).  Per-level observation times are \texttt{launch\_time} plus \texttt{ETIME} at each level, but only ${\sim}15\%$ of profiles in the current subsample carry any valid \texttt{ETIME} values; for the remaining ${\sim}85\%$ every \texttt{observation\_time} bin is left missing.  To provide a usable per-level time coordinate for the \texttt{ETIME}-missing majority, IGRA output files carry an additional coordinate \texttt{estimated\_observation\_time}: equal to \texttt{observation\_time} where the latter is finite, and otherwise synthesised under a piecewise-constant ascent model, $\SI{5.0}{m/s}$ at or below $\SI{20}{km}$ and $\SI{6.0}{m/s}$ above.  The synthesis is anchored to remain continuous with any reported \texttt{ETIME}: if a profile has at least one finite \texttt{observation\_time}, the synthesis anchor is the first such level's $(z, t)$; otherwise the anchor is $(z_{\text{station}}, \text{launch\_time})$ using the station elevation from the IGRA metadata catalog.  For the ${\sim}35\%$ of station files that have no metadata-catalog entry, the fallback anchor is $(z_{\text{lowest finite}}, \text{launch\_time})$ in the profile (placing zero elapsed time at the first observed level rather than at the true surface).  The two-zone rate is calibrated from the $\sim\!217{,}000$ profiles with valid \texttt{ETIME}, which give a median local ascent rate of $\SI{5.0}{m/s}$ from 1--$\SI{20}{km}$ and $\SI{5.3}$--$\SI{6.7}{m/s}$ in the 20--$\SI{30}{km}$ band as the balloon encounters lower-density air and expands.  The single kink at $\SI{20}{km}$ is a deliberate simplification: with $\SI{6.0}{m/s}$ as the constant above-$\SI{20}{km}$ rate, the integrated time agrees with the height-varying true rate to within ${\sim}60\,\mathrm{s}$ at $\SI{30}{km}$, which is small compared to the $\SI{10}$--$\SI{100}{s}$ bin-crossing time of a single $\SI{100}{m}$ grid cell.  Consumers needing a per-level time for structure functions or compositing may use \texttt{estimated\_observation\_time} directly; \texttt{observation\_time} remains as the reported (where available) quantity.  For IGRA the horizontal drift track (\S3.1) is also integrated using \texttt{estimated\_observation\_time} rather than \texttt{observation\_time}, so drift tracks are defined for the full archive rather than only the ${\sim}15\%$ of profiles with reported \texttt{ETIME}.

\subsection{JOANNE}

The Joint dropsonde Observations of the Atmosphere in tropical North atlaNtic meso-scale Environments (JOANNE) dataset \citep{george2021} contains 1215 dropsondes deployed during the EUREC\textsuperscript{4}A field campaign in January--February 2020.  Dropsondes were released from the German HALO and NOAA WP-3D aircraft over the tropical North Atlantic east of Barbados, with the goal of characterizing the thermodynamic and dynamic environment of trade-wind cumulus clouds.  Sondes were deployed along circular flight patterns (${\sim}\SI{222}{km}$ diameter) to enable estimation of area-averaged mesoscale divergence, vorticity, and vertical velocity.  The dataset provides both individual sounding profiles and a uniformly gridded product at $\SI{10}{m}$ vertical spacing.  \citet{george2021} assign each profile a quality control flag of ``good,'' ``bad,'' or ``ugly'' based on launch detection, profile completeness ($\geq$80\% for all variables), and plausibility of near-surface values.  A multiplicative correction factor of 1.06 was applied to relative humidity from HALO sondes to compensate for a systematic dry bias caused by insufficient reconditioning of the humidity sensor before flight.

Here, only profiles flagged as ``good'' are used.

\subsection{BEACH}

The Barbados and Eastern Atlantic Combined High-altitude (BEACH) dropsonde dataset \citep{gloeckner2025} contains 1191 dropsondes deployed from the HALO aircraft during the ORCESTRA field campaign in August--September 2024.  Soundings were coordinated by the PERCUSION and MAESTRO subcampaigns and targeted the structure and variability of the Atlantic Inter-tropical Convergence Zone (ITCZ).  Circular flight patterns at approximately $\SI{14}{km}$ altitude (${\sim}\SI{260}{km}$ diameter) were flown over both the eastern and western tropical Atlantic, with flight tracks designed to coincide with EarthCARE satellite overpasses and to sample within and across the ITCZ.  Like JOANNE, the dataset includes a gridded product at $\SI{10}{m}$ spacing and derived mesoscale circulation diagnostics.  \citet{gloeckner2025} assign each profile an aggregate quality flag (GOOD, BAD, or UGLY) based on profile sparsity, profile extent above $\SI{8000}{m}$, near-surface coverage, and physical plausibility of surface values.  Seventeen sondes were falsely configured by the manufacturer as minisondes rather than NRD41 (listed in supplement Table~E2 of \citealt{gloeckner2025}), and one sonde (0bd0e322, flight HALO-20240924a) had a factory-reset sensor module producing anomalous temperature and humidity.

Here, only profiles flagged as ``GOOD'' are used, and the 17 misconfigured and 1 factory-reset sondes listed above are excluded.

\subsection{OTREC}

The Organization of Tropical East Pacific Convection (OTREC) dropsonde dataset \citep{vomel2021} contains 648 dropsondes released from the NSF/NCAR Gulfstream-V during 22 research flights between August and October 2019.  The campaign investigated the distribution and day-to-day variability of deep atmospheric convection in the tropical eastern Pacific and Caribbean, with a focus on why higher rainfall rates occur over lower sea surface temperatures in this region.  Soundings were launched in grid patterns with a typical horizontal spacing of $1^\circ$ longitude and $1.2^\circ$ latitude, from a mean flight altitude of $\SI{13.3}{km}$ to the surface.  OTREC was the first field campaign to rely entirely on the NCAR NRD41 dropsonde, with a nominal vertical resolution between 6 and $\SI{12}{m}$.  All soundings in the distributed dataset have passed ASPEN quality control with manual post-processing corrections; erroneous data points are set to missing.  \citet{vomel2021} applied a per-sonde pressure offset correction (median $\SI{0.35}{hPa}$) and a relative humidity time-lag correction for the slow sensor response above ${\sim}\SI{11.5}{km}$.  One sounding (20190925\_154412) exhibits an unexplained warm bias of 1--$\SI{3.7}{K}$ throughout the profile.

Here, the corrected dataset is used as provided, with the exception of the warm-biased sounding noted above, which is excluded.

\subsection{ACTIVATE}

The Aerosol Cloud meTeorology Interactions oVer the western ATlantic Experiment (ACTIVATE) dropsonde dataset \citep{vomel2023activate} contains 801 dropsondes released from NASA Langley's King Air research aircraft during 165 flights between February 2020 and June 2022.  The campaign studied aerosol--cloud interactions over the western North Atlantic, off the coast of the eastern United States and around Bermuda, sampling all seasons across six deployment periods.  The King Air flew at approximately $\SI{9}{km}$ altitude, and resulting dropsonde profiles provide temperature, pressure, relative humidity, and horizontal and vertical winds between the surface and flight level.  \citet{vomel2023activate} processed all soundings through ASPEN and applied corrected launch times for 17 sondes affected by a synchronization error between the onboard computer and GPS time.  Twenty sondes with unreconditioned humidity sensors may exhibit a dry bias.

Here, profiles marked as degraded are excluded.  The corrected launch times are used where applicable, and only temperature, pressure, and wind from the 20 sondes with unreconditioned humidity sensors is retained.

\subsection{Hurricane dropsonde dataset}

The NOAA hurricane GPS dropsonde dataset \citep{wang2015} consists of 13,681 atmospheric profiles collected during NOAA hurricane reconnaissance and surveillance flights between 1996 and 2012, covering 120 tropical cyclones.  Dropsondes were released from NOAA Gulfstream-IV and WP-3D aircraft in the inner and outer core regions of tropical cyclones, typically at $\SI{150}{}$--$\SI{200}{km}$ horizontal intervals.  The dataset uses the NCAR Airborne Vertical Atmospheric Profiling System (AVAPS), which underwent a major redesign in 2008 (AVAPS II) that expanded simultaneous sounding capacity to eight channels and improved GPS wind recovery rates.  \citet{wang2015} processed all profiles through ASPEN and performed visual inspection; the distributed dataset constitutes a single quality-controlled product.

Here, the archive is used as provided.

\subsection{DYNAMO}

The Dynamics of the Madden--Julian Oscillation (DYNAMO) field campaign \citep{ciesielski2014} investigated the initiation and propagation of the MJO over the tropical Indian Ocean during Fall 2011 through Spring 2012.  In addition to the campaign's extensive radiosonde network, 470 GPS dropsondes were released from the NOAA P-3 aircraft during 13 research flights between 9~November and 13~December 2011.  The P-3 was based at Diego Garcia and flew at an average altitude of $\SI{2}{km}$ (maximum ${\sim}\SI{8}{km}$), with missions sampling the central Indian Ocean ($14.5^\circ$S--$1.2^\circ$N, $69.8^\circ$--$80.9^\circ$E) for convective systems, boundary layer structure, and ITCZ variability.  Of the 470 deployed, 469 passed quality control and are included in the final dropsonde archive.  All profiles were processed through ASPEN; the Version~3 release incorporates a temperature-dependent dry bias correction to relative humidity that was applied to all RD94 dropsondes collected from 2010 onward.

Here, 468 profiles from the dropsonde archive are used as provided.

\subsection{SHOUT}

The Sensing Hazards with Operational Unmanned Technology (SHOUT) project \citep{wick2020} deployed 638 GPS dropsondes from NASA's Global Hawk unmanned aircraft (cruise altitude ${\sim}\SI{18}{km}$) during three field campaigns in 2015 and 2016.  The Global Hawk flew 15 missions sampling 6 tropical cyclones and 3 Pacific winter storms, with missions designed to target regions of high model forecast uncertainty identified through adjoint and ensemble sensitivity analyses.  All profiles were processed through ASPEN.

We use the archive as provided.

\subsection{AR Recon}

Atmospheric River Reconnaissance (AR Recon) is a recurring campaign that has deployed dropsondes from NOAA G-IVSP and Air Force Reserve WC-130J aircraft during winter seasons (November--March) since 2016, targeting atmospheric rivers approaching the U.S.\ West Coast with the primary aim of improving forecasts of landfalling events.  \citet{cobb2022} describe the 2021 season, during which 1142 dropsondes were released during 45 flights.  The data used here span the 2018--2026 winter seasons.  Drop locations are guided by adjoint and ensemble sensitivity analyses to target regions where additional observations would most reduce forecast uncertainty.  All profiles were processed through ASPEN quality control.

Here, the archive is used as provided.

\subsection{ENRR}

The NOAA El Ni\~no Rapid Response (ENRR) field campaign \citep{dole2018} deployed dropsondes from three aircraft platforms during the strong 2015/16 El Ni\~no event, from January through March 2016.  The NOAA Gulfstream-IV (G-IV) flew 22 missions releasing 607 dropsondes over the central tropical Pacific, targeting deep convective regions and tropical--extratropical linkages.  A NOAA C-130 flew 6 missions releasing 280 dropsondes targeting atmospheric rivers between Hawaii and the U.S.\ West Coast, and the NASA Global Hawk flew 3 missions releasing 89 dropsondes from cruise altitude (${\sim}\SI{18}{km}$).  All profiles were processed through ASPEN; the G-IV and Global Hawk archives apply corrected data, and the C-130 archive applies a temperature-dependent dry bias correction.

Here, the corrected archives from all three platforms are used as provided (976 profiles total).

\subsection{HS3}

NASA's Hurricane and Severe Storm Sentinel (HS3) investigation \citep{braun2016} deployed dropsondes from the NASA Global Hawk unmanned aircraft (cruise altitude ${\sim}\SI{18}{km}$) during the Atlantic hurricane seasons of 2012--2014.  The ``environmental'' Global Hawk (AV-6) carried instruments to characterize the storm environment, including dropsondes for temperature, humidity, and wind profiles.  HS3 flew 21 missions over 9 named storms, along with flights over Saharan Air Layer outbreaks and non-developing systems.  A 2011 precursor deployment contributed two intercomparison and atmospheric river sampling flights.  All profiles were processed through ASPEN.  The dropsonde archive contains 1505 quality-controlled soundings across the four years: 79 in 2011, 337 in 2012, 433 in 2013, and 656 in 2014.

Here, all 1505 quality-controlled profiles are used as provided.

\subsection{PREDICT}

The Pre-Depression Investigation of Cloud-systems in the Tropics (PREDICT) experiment \citep{montgomery2012} deployed 568 dropsondes from the NSF/NCAR Gulfstream-V during 26 flights over the Caribbean and western Atlantic in August--September 2010.  The campaign investigated tropical cyclogenesis, with priority given to developing systems prior to their classification as tropical depressions.  Of the 568 deployed, 558 passed quality control and are included in the final archive.  All profiles were processed through ASPEN.  The Version~3 release incorporates a temperature-dependent relative humidity and dewpoint dry bias correction.

Here, the 558 quality-controlled profiles are used as provided.

\subsection{HALO--(AC)\textsuperscript{3} dropsondes}

The HALO--(AC)\textsuperscript{3} aircraft campaign \citep{ehrlich2025} collected in situ and remote sensing measurements over the Norwegian and Greenland Seas, the Fram Strait, and the central Arctic Ocean in March--April 2022.  The HALO and Polar~5 aircraft deployed dropsondes during 30 flight segments, with profiles spanning altitudes from the surface to ${\sim}\SI{13}{km}$.  \citet{george2024} processed the dropsonde data through quality control and provide 464 Level-2 individual sounding profiles as NetCDF files, with units and structure similar to JOANNE and BEACH.

Here, 463 of the 464 Level-2 profiles are used.  One sonde (163313035, deployed 2022-04-07) is excluded because its GPS receiver failed, producing no altitude data.

\section{Regridding algorithm}

The output grid is a one-dimensional partition of altitude into cells of uniform thickness $\Delta z$.  Cell edges are placed at
\begin{align}
    z_k = k\,\Delta z, \qquad k = 0, 1, \ldots, N, \label{eq:edges}
\end{align}
where $N = \lfloor z_{\max}/\Delta z\rfloor$ is the number of cells.  The $k$-th cell spans the half-open interval $[z_k,\, z_{k+1})$.  Cell-center altitudes, used as the coordinate for all output variables, are
\begin{align}
    \bar{z}_k = (k + \tfrac{1}{2})\,\Delta z. \label{eq:centers}
\end{align}
The grid spacing is set to $\Delta z = \SI{10}{m}$ for dropsonde datasets, chosen to be finer than the effective vertical resolution of most modern dropsondes (of order $\SI{200}{m}$) so that no information is lost during regridding.  IGRA radiosondes use a coarser spacing of $\Delta z = \SI{100}{m}$ to reduce file sizes given the large number of stations and soundings.  The grid extends to $z_{\max} = \SI{20}{km}$ for all dropsonde datasets and $z_{\max} = \SI{40}{km}$ for IGRA radiosondes, which routinely ascend into the stratosphere.

For each grid cell, the regridded value of a variable is equal to the arithmetic mean of all measurements whose altitude falls within that cell.  If a cell contains no measurements, the value is recorded as missing, and this procedure is applied independently to each variable.

All dropsonde datasets report geometric (GPS-derived) altitude.  IGRA reports geopotential height, which is converted to geometric altitude via $z = R_e H / (R_e - H)$, where $H$ is geopotential height and $R_e = \SI{6371}{km}$ is the mean Earth radius.  The difference is approximately 0.5\% at $\SI{30}{km}$ (${\sim}\SI{100}{m}$).

The following variables are bin-averaged onto the uniform grid (Tab. \ref{tab:variables}): zonal and meridional wind ($u$, $v$), pressure ($p$), temperature ($T$), and relative humidity ($\mathrm{RH}$).  When a dataset provides the water vapor mixing ratio $q$ directly rather than relative humidity, $q$ is bin-averaged in place of $\mathrm{RH}$ and no subsequent diagnosis is needed. 


Five additional variables are diagnosed from the regridded fields (Tab. \ref{tab:variables}).  Diagnosis is performed after bin averaging so that the thermodynamic relations are evaluated on the gridded values rather than on individual high-frequency measurements. 

\begin{table}[h]
\centering
\caption{Output variables.  Reported variables are bin-averaged directly; diagnosed variables are computed from the gridded fields after bin averaging.}
\label{tab:variables}
\begin{tabular}{lllll}
\toprule
Variable & Description & Units & Type \\
\midrule
\texttt{u}                  & Zonal wind component             & \si{m\per s}    & Reported \\
\texttt{v}                  & Meridional wind component         & \si{m\per s}    & Reported \\
\texttt{p}                  & Pressure                          & \si{Pa}         & Reported \\
\texttt{T}                  & Temperature                       & \si{K}          & Reported \\
\texttt{RH}                 & Relative humidity                 & \%              & Reported \\
\texttt{observation\_time}  & Measurement time at each level    & datetime        & Reported \\
\texttt{q}                  & Water vapor mixing ratio          & \si{kg\per kg}  & Diagnosed (Eq.~\ref{eq:q}) \\
\texttt{theta}              & Potential temperature              & \si{K}          & Diagnosed (Eq.~\ref{eq:theta}) \\
\texttt{theta\_e}           & Equivalent potential temperature   & \si{K}          & Diagnosed (Eq.~\ref{eq:theta_e}) \\
\texttt{MSE}                & Moist static energy               & \si{J\per kg}   & Diagnosed (Eq.~\ref{eq:MSE}) \\
\texttt{DSE}                & Dry static energy                 & \si{J\per kg}   & Diagnosed (Eq.~\ref{eq:DSE}) \\
\texttt{x\_offset}          & Eastward offset from launch        & \si{m}          & Diagnosed (\S\ref{sec:drift}) \\
\texttt{y\_offset}          & Northward offset from launch       & \si{m}          & Diagnosed (\S\ref{sec:drift}) \\
\texttt{lat}                & Latitude at each level             & \si{\degree}N   & Diagnosed (\S\ref{sec:drift}) \\
\texttt{lon}                & Longitude at each level            & \si{\degree}E   & Diagnosed (\S\ref{sec:drift}) \\
\texttt{u\_clim}            & ERA5 climatological zonal wind     & \si{m\per s}    & Attached (\S\ref{sec:clim}) \\
\texttt{v\_clim}            & ERA5 climatological meridional wind & \si{m\per s}   & Attached (\S\ref{sec:clim}) \\
\bottomrule
\end{tabular}
\end{table}

When relative humidity is the reported moisture variable, the mixing ratio is computed from the regridded $\mathrm{RH}$, $T$, and $p$ fields.  The saturation vapor pressure over liquid water $e_s$ is obtained from the August--Roche--Magnus formula,
\begin{align}
    e_s(T) = 610.94 \exp\!\left(\frac{17.625\, T_c}{T_c + 243.04}\right), \label{eq:es}
\end{align}
where $T_c = T - \SI{273.15}{K}$ is the temperature in degrees Celsius and $e_s$ has units of Pascals.  The vapor pressure is $e = (\mathrm{RH}/100)\, e_s$, and the mixing ratio follows as
\begin{align}
    q = \frac{0.622\, e}{p - e}. \label{eq:q}
\end{align}


Potential temperature is computed from Poisson's equation,
\begin{align}
    \theta = T \left(\frac{p_0}{p}\right)^{\kappa}, \label{eq:theta}
\end{align}
where $p_0 = \SI{100000}{Pa}$ is the reference pressure and $\kappa = R_d / c_p \approx 0.2854$ is the Poisson constant, with $R_d = \SI{287.04}{J\,kg^{-1}\,K^{-1}}$ the specific gas constant for dry air and $c_p = \SI{1005.7}{J\,kg^{-1}\,K^{-1}}$ the isobaric specific heat.


Moist static energy combines sensible heat, gravitational potential energy, and latent heat into a single conserved quantity for moist adiabatic processes:
\begin{align}
    h = c_p T + g\bar{z}_k + L_v q, \label{eq:MSE}
\end{align}
where $g = \SI{9.807}{m\,s^{-2}}$ is gravitational acceleration, $L_v = \SI{2.501e6}{J\,kg^{-1}}$ is the latent heat of vaporization at $\SI{0}{\celsius}$, and $\bar{z}_k$ is the cell-center altitude from Eq.~\ref{eq:centers}.


Dry static energy is conserved for dry adiabatic processes:
\begin{align}
    s = c_p T + g\bar{z}_k. \label{eq:DSE}
\end{align}


Equivalent potential temperature is conserved for saturated adiabatic processes.  It is approximated following \citet{bolton1980}:
\begin{align}
    \theta_e = \theta \exp\!\left(\frac{L_v\, q}{c_p\, T_L}\right), \label{eq:theta_e}
\end{align}
where $T_L$ is the temperature at the lifting condensation level, estimated as
\begin{align}
    T_L = \frac{1}{\displaystyle\frac{1}{T - 55} - \frac{\ln(\mathrm{RH}/100)}{2840}} + 55. \label{eq:TL}
\end{align}

\subsection{Horizontal drift integration} \label{sec:drift}

Most source datasets report per-level GPS coordinates, so that the horizontal position of the sonde is known at each altitude level as it drifts with the wind during its flight.  A small fraction of older files (notably 306 of 13{,}678 files in the 1996--2012 hurricane archive) lack position data entirely, and IGRA radiosondes never carry per-level coordinates.  Rather than retain the native coordinates for some datasets and not others, we compute the horizontal drift track uniformly for every profile by integrating the gridded horizontal winds in time, starting from the launch position. Datasets that do report per-level coordinates are used as validation (\S\ref{sec:validation}).

For a single profile, let $\{\bar{z}_k\}$ denote the cell-center altitudes and $\{t_k\}$ the bin-averaged observation times.  The altitude bins are reindexed in ascending order of $t_k$ such that dropsondes progress downward and radiosondes upward.  A horizontal position offset, representing horizontal drift, is accumulated from the initial launch position using the trapezoidal rule on the gridded winds:
\begin{align}
    \Delta x_{i \to i+1} &= \tfrac{1}{2}\left(u_i + u_{i+1}\right)\left(t_{i+1} - t_i\right), \label{eq:dx} \\
    \Delta y_{i \to i+1} &= \tfrac{1}{2}\left(v_i + v_{i+1}\right)\left(t_{i+1} - t_i\right), \label{eq:dy}
\end{align}
where $i$ indexes the time-sorted bins.  The cumulative east- and north-offsets from the launch point are
\begin{align}
    x_k = \sum_{i=0}^{k-1} \Delta x_{i \to i+1}, \qquad y_k = \sum_{i=0}^{k-1} \Delta y_{i \to i+1}.
\end{align}
Bins whose observation time is missing are omitted from the traversal and their offsets are recorded as missing.

When $u$ or $v$ is missing at a bin, the integration carries forward the last valid value across the gap. If the gap becomes larger than one kilometer, corresponding to 100 dropsonde bins or 10 IGRA bins, the remaining offset positions are marked as missing for the remainder of the profile. This threshold limits the horizontal error that can be incurred by extrapolating through a single unobserved region.

A subset of profiles in some campaign archives exhibit a large disagreement between $\int u\,dt$ and the sonde's own GPS track, which we attribute to inaccuracies in the provider's ASPEN fall-speed model for those particular sondes.  These profiles would otherwise carry a horizontal drift track that does not reflect where the sonde actually was.  We screen for them after the drift integration by comparing the gridded surface drift to the native GPS drift at the lowest altitude bin with valid GPS; profiles whose disagreement exceeds \SI{2}{km} have their \texttt{x\_offset}, \texttt{y\_offset}, \texttt{lat}, and \texttt{lon} fields set to missing, while all other variables are retained.  The screening is implemented in \texttt{src/screen\_drift.py} and runs as a post-processing step.

Latitude and longitude at each level are obtained from the east- and north-offsets using a small-angle spherical correction anchored at the launch latitude:
\begin{align}
    \phi_k &= \phi_0 + \frac{180^\circ}{\pi}\frac{y_k}{R_e}, \label{eq:lat} \\
    \lambda_k &= \lambda_0 + \frac{180^\circ}{\pi}\frac{x_k}{R_e \cos\phi_0}, \label{eq:lon}
\end{align}
where $\phi_0$ and $\lambda_0$ are the launch latitude and longitude, respectively, and $R_e = \SI{6371}{km}$ is the mean Earth radius.  For profiles without a launch location (e.g.\ the synthetic SAM and multifractal datasets), $x_k$ and $y_k$ are still computed when winds and times are available, but $\phi_k$ and $\lambda_k$ are left missing.

\subsection{Launch position} \label{sec:launch}

For every dataset whose source files report per-level GPS coordinates, the stored \texttt{launch\_lat}/\texttt{launch\_lon} is taken from the earliest-time per-level GPS fix rather than from provider metadata.  This is a deliberate departure from the convention of some providers.  The raison d'\^etre is that several archives populate a launch-position attribute from an aircraft centroid or mission plan that does not exactly match the sonde's own GPS stream.  For HALO--(AC)\textsuperscript{3} we observed offsets of up to ${\sim}\SI{15}{km}$ between the \texttt{aircraft\_latitude\_(deg\_N)} attribute and the first per-level GPS fix; offsets of several hundred metres are routine for other campaigns.  Using the sonde's own GPS removes this dependency on provider metadata, and ensures that the launch position is consistent with the horizontal drift track integrated in \S\ref{sec:drift}.

For IGRA radiosondes, per-level GPS is unavailable and the launch position is the station location from the IGRA metadata catalog.  For the synthetic SAM datasets the simulation domain is periodic and not georeferenced, so \texttt{launch\_lat}/\texttt{launch\_lon} are missing; a launch position is still recorded as \texttt{launch\_x}/\texttt{launch\_y} in LES-domain metres.  The multifractal datasets have no spatial context, so both forms of launch position are missing.

\subsection{Climatological wind profiles} \label{sec:clim}

A monthly climatological profile of zonal and meridional wind is attached to every sounding as a reference against which the instantaneous measurements can be assessed.  The climatology is derived by averaging ERA5 monthly mean winds and geopotential height over the period 1995--2025. The pressure coordinate is interpolated to a \SI{100}{m} geopotential height grid spanning 0 to \SI{40}{km}, retaining the $0.25^\circ\times 0.25^\circ$ horizontal resolution. Values below the ERA5 surface orography are stored as missing.

For each launch location $(\phi_0, \lambda_0)$ and calendar month $m$, the climatological wind is obtained by (i) horizontal bilinear interpolation of the regridded ERA5 dataset for month $m$ onto $(\phi_0, \lambda_0)$, then (ii) linear interpolation of the resulting column in altitude onto the output cell centers $\bar{z}_k$.  Target altitudes above the ERA5 ceiling or below the local ERA5 surface orography are marked as missing.


The storage layout differs between dataset types.  For dropsonde campaigns, each sounding has its own launch location and time, and the climatology for the relevant month is stored per sounding; the variables \texttt{u\_clim} and \texttt{v\_clim} share the \texttt{(sounding\_id, altitude)} shape of the directly gridded fields.  For IGRA station files, the station location is fixed but individual soundings span many different months, so instead all twelve monthly profiles are stored for each the station location with shape \texttt{(month, altitude)}, where \texttt{month} is a new coordinate indexing calendar months $1\dots 12$.  Profiles for which a launch location or launch time is not available are stored as missing, which includes the synthetic SAM and multifractal datasets.

\subsection{Output format} \label{sec:output}

The regridded data are stored as one NetCDF-4 file per dropsonde campaign, or one file per station for IGRA (in an \texttt{igra/} subdirectory).  Each file has two dimensions: \texttt{sounding\_id} and \texttt{altitude}.  The \texttt{sounding\_id} dimension indexes individual profiles.  For dropsonde datasets each profile corresponds to a unique spatial location; for IGRA station files all profiles share the same station location but differ in launch time.
The \texttt{altitude} coordinate contains the cell-center altitudes $\bar{z}_k$ from Eq.~\ref{eq:centers}, in meters above mean sea level.

Each launch location carries the following metadata:

\begin{center}
\begin{tabular}{lll}
\toprule
Variable & Type & Description \\
\midrule
\texttt{launch\_time}        & datetime & Date and time of sounding launch (\texttt{sounding\_id}) \\
\texttt{observation\_time}  & datetime & Observation time at each altitude level (\texttt{sounding\_id}, \texttt{altitude}) \\
\texttt{launch\_lat}        & float    & Latitude at profile start [\si{\degree}N] (\texttt{sounding\_id}) \\
\texttt{launch\_lon}        & float    & Longitude at profile start [\si{\degree}E] (\texttt{sounding\_id}) \\
\texttt{sonde\_id}          & string   & Sonde or profile identifier (\texttt{sounding\_id}) \\
\bottomrule
\end{tabular}
\end{center}

\noindent The \texttt{observation\_time} variable records the actual measurement time at each altitude level, bin-averaged within each grid cell.  For dropsonde datasets, \texttt{launch\_lat} and \texttt{launch\_lon} refer to the aircraft position at the time of release, or the first GPS measurement in the profile when explicit release coordinates are not available.  For IGRA radiosondes, they refer to the launch site; the station identity is encoded in the filename and in the \texttt{sonde\_id} (format: \texttt{\{station\_id\}\_\{YYYYMMDDHH\}}).

\subsection{Data variables}

All regridded and diagnosed variables are stored with shape (\texttt{sounding\_id}, \texttt{altitude}).  Missing values (grid cells with no measurements) are stored as NaN.



\section{Validation} \label{sec:validation}

Two of the dropsonde datasets---JOANNE \citep{george2021} and BEACH \citep{gloeckner2025}---provide both individual sounding profiles and a uniformly gridded product at $\SI{10}{m}$ vertical spacing produced by the data providers. 
Agreement between the two gridded products is expected to be close but not exact given that the provider gridded products use interpolation rather than bin averaging, and in the case of BEACH use slightly different filtering criteria.

For each variable, the comparison is performed per sounding at all grid cells where both products report non-missing values.  Across all cells, the median RMSD and mean bias between our regridded product and the official products are negligible (Tab. \ref{tab:validation}), with the possible exception of the BEACH pressure variable. A small discrepancy is to be expected here given that the gridded pressure variable from \citet{gloeckner2025} obtained pressure through integrating the thermodynamic profile, not interpolating measured pressure values.

\begin{table}[h]
\centering
\caption{Validation statistics comparing bin-averaged profiles to provider-gridded products.  RMSD and bias are computed per sounding over matched non-missing grid cells; the table reports the median RMSD and mean bias across all matched sondes.}
\label{tab:validation}
\begin{tabular}{llrrr}
\toprule
Dataset & Variable & Median RMSD & Mean bias & Units \\
\midrule
JOANNE (1068 sondes) & $T$             & 0.012  & $<$0.001      & \si{K} \\
                      & $u$             & 0.025  & $-$0.000      & \si{m\per s} \\
                      & $v$             & 0.024  & $<$0.001      & \si{m\per s} \\
                      & $\mathrm{RH}$   & 0.12   & $<$0.001      & \% \\
                      & $p$             & 11.3   & $-$0.003      & \si{Pa} \\
\midrule
BEACH (976 sondes)    & $T$             & 0.24   & $-$0.18       & \si{K} \\
                      & $u$             & 0.27   & $-$0.023      & \si{m\per s} \\
                      & $v$             & 0.26   & $<$0.001      & \si{m\per s} \\
                      & $\mathrm{RH}$   & 1.65   & $-$0.091      & \% \\
                      & $p$             & 137    & $-$120        & \si{Pa} \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Horizontal drift}

The drift integration described in \S\ref{sec:drift} is checked against the per-level GPS coordinates reported in the source files, for each dataset whose raw data contains them.  For a given dataset, a random subset of 20 profiles is re-read from source and up to 10 native (altitude, latitude, longitude) triples are drawn per profile.  The native samples and the gridded output are expressed as east/north offsets from the earliest-time sample in each stream, so that any uniform anchor offset per profile cancels and only integration error remains.  The test reports the maximum and RMS of these residual offsets across all sampled levels; values are summarized in Tab.~\ref{tab:drift-validation}.

\begin{table}[h]
\centering
\caption{Drift validation statistics.  ``Flagged'' is the number of profiles whose integrated drift disagreed with native GPS by more than $\SI{2}{km}$ at the lowest altitude bin with a valid integration (\S\ref{sec:drift}) and whose \texttt{x\_offset}, \texttt{y\_offset}, \texttt{lat}, \texttt{lon} have therefore been set to missing (the rest of the regridded and diagnosed fields are retained).  ``Max residual'' and ``RMS residual'' are the maximum and RMS of $|\mathbf{d}_\text{grid} - \mathbf{d}_\text{native}|$ over a random sample of 20 unflagged profiles $\times$ up to 10 levels each, where each drift vector $\mathbf{d}$ is expressed as an east/north offset from its own earliest-time GPS fix so that any uniform per-profile anchor offset cancels.  Flagged profiles are excluded from these statistics because their drift fields are entirely missing.  Launch residual is the distance between the stored \texttt{launch\_lat}/\texttt{launch\_lon} and the earliest-time per-level GPS fix and is zero by construction of \S\ref{sec:launch}.}
\label{tab:drift-validation}
\begin{tabular}{lrrrrr}
\toprule
Dataset           & Profiles  & Flagged        & Max residual [m]  & RMS residual [m]  & Launch residual [m] \\
\midrule
JOANNE            & 1068      & 0              & 234             & 126             & 0 \\
BEACH             & 976       & 0              & 284             & 134             & 0 \\
OTREC             & 646       & 0              & 500             & 272             & 0 \\
ACTIVATE          & 782       & 0              & 730             & 354             & 0 \\
Hurricane         & 13{,}647  & 203            & 542             & 175             & 0 \\
DYNAMO            & 468       & 0              & 278             & 123             & 0 \\
SHOUT             & 634       & 2              & 1120            & 755             & 0 \\
AR~Recon          & 8377      & 5              & 1540            & 282             & 0 \\
ENRR              & 965       & 0              & 948             & 221             & 0 \\
HS3               & 1505      & 75             & 1240            & 722             & 0 \\
PREDICT           & 555       & 2              & 586             & 223             & 0 \\
HALO--(AC)\textsuperscript{3} & 463 & 0          & 702             & 214             & 0 \\
\bottomrule
\end{tabular}
\end{table}

The residuals that remain after flagging are dominated by small mismatches between $\int u\,dt$ and the sonde's own GPS track that are inherent to the source data, not introduced by the regridding.  The effect is most visible in SHOUT and HS3, where per-profile RMS residuals of $\sim\SI{0.7}{km}$ and maxima exceeding $\SI{1}{km}$ are normal.  JOANNE, BEACH, and DYNAMO are notably cleaner, with maximum errors under $\SI{300}{m}$.  The launch residual of zero everywhere is a consequence of the launch convention in \S\ref{sec:launch}, not an independent check on it.

IGRA radiosondes and the synthetic SAM and multifractal datasets are not included in this validation because their source files do not carry per-level GPS coordinates against which the integrated drift could be compared.  The drift fields \texttt{x\_offset}, \texttt{y\_offset}, and (for IGRA) \texttt{lat}/\texttt{lon} are still computed for them from the gridded $u$ and $v$ as described in \S\ref{sec:drift}; the flagging step in \texttt{src/screen\_drift.py} is simply a no-op for these datasets.

\bibliographystyle{apalike}
\bibliography{references}

\end{document}
